---
title: "JSON and Maps"
author: "Pablo Ugarte"
date: '2017-09-10'
slug: json-and-maps
tags: []
categories:
- json
- jsonlite
- Heatmap
- leaflet
---

Today we will extract information from a JSON document using the R package jsonlite and we will then plot lon/lat in a map using the R package leaflet.

The data that we will be looking at is crime in LA.

But first, we need to install and load libraries for this excercise.

```{r, message = FALSE, warning = FALSE}
library(jsonlite)
library(tidyverse)
library(stringr)
library(tibble)
library(leaflet)
library(ggmap)
library(readr)
```

Now we can start grabbing the data from the json document which can be found here [^1].

```{r, message = FALSE, warning = FALSE}
crimeLARaw <- fromJSON("https://data.lacity.org/resource/7fvc-faax.json")
```

We will now clean the data by just selecting certain columns.

```{r, message = FALSE, warning = FALSE}
crimeLARaw_tbl <- crimeLARaw %>% 
  select(crm_cd_desc , premis_desc, time_occ, location_1)
```

I am having problems extracting the long/lat data without parentheses. So decided to take this dataframe, and copy and paste to an excel file. 
In excel, with the find and replace function we can get easily rid of the "c(" and ")". We then save it as csv. 
We can now return to R and continue data wrangling.

```{r, message = FALSE, warning = FALSE}
crimeLA <- read_csv("crimeLA.csv")
```

We split the coordinates column and convert variables to numeric and factor format.

```{r, message = FALSE, warning = FALSE}
crimeLA <- crimeLA %>% separate(cordinates, c("long", "lat"), sep = ", ")
crimeLA$long <- as.numeric(crimeLA$long) 
crimeLA$lat <- as.numeric(crimeLA$lat)
crimeLA$crime_type <- as.factor(crimeLA$crime_type)
```

Last, we plot the long/lat info. We also create a palette that maps factor levels to colors.

```{r, echo = TRUE, message = FALSE, warning = FALSE}
pal <- colorFactor(c("gray","red","green3","blue","cyan","magenta","yellow","black"),
                   domain = c("BURGLARY", "BURGLARY FROM VEHICLE","CHILD ANNOYING (17YRS & UNDER","CRUELTY TO ANIMALS","DISTURBING THE PEACE","EMBEZZLEMENT, GRAND THEFT ($950.01 & OVER","INDECENT EXPOSURE","OTHER ASSAULT"  ,"OTHER MISCELLANEOUS CRIME","THEFT PLAIN - PETTY ($950 & UNDER","VANDALISM - MISDEAMEANOR ($399 OR UNDER","VEHICLE - STOLEN" ))

leaflet(data = crimeLA) %>% 
  addTiles() %>%
  addCircles(~long, ~lat) %>% 
  addMarkers(popup = ~as.character(crime_type), label = ~as.character(crime_type)) %>%
  addCircleMarkers(color = ~pal(crime_type)) %>%
  addLegend(pal = pal, values = ~crime_type, opacity = 5)
```

Thats it. We now have a map that shows crime in LA. Happy blog!

[^1]: [Link Youtube] (https://catalog.data.gov/dataset/crime-data-from-2010-to-present)

