---
title: "Tennis winners - scraping and plotting"
author: "Pablo"
date: '2017-09-11'
slug: tennis-winners-scrapig-and-plotting
tags: []
categories:
- scraping
- ggtheme
---

# Intro

>In this tutorial, we will be plotting tennis players Grand Slam Tournament wins. 

The data comes from the Tennis Grand Slam Tournaments results from the ESP website [^1].

We will scrape the data and continue with the visualization thereafter.

# Libraries

Overall, we are going to use the following R packages:

```{r, message = FALSE, warning = FALSE}
library(XML) # for web scraping
library(rvest) # for web scraping
library(dplyr) # for dataset manipulation
library(knitr) # for nice dataset printing
library(hrbrthemes) # Aesthetic defaults for ggplot2 charts
library(ggthemes) # Themes ggplot2
library(ggplot2) # Charts
library(stringr) # Manipulating strings
```

# Scraping

We will be scraping HTML Table with the XML package. By passing the URL to readHTMLTable() we extract data from HTML tables in HTML documents. The data in each table is read and stored as a data frame. We can use the which argument in readHTMLTable() to import only those tables specified numerically.

```{r, message = FALSE, warning = FALSE}
url <- "http://www.espn.com/tennis/history"

# Read in HTML data only parsing the 1st table
tbls_xml <- readHTMLTable(url, which = 1)
```

# Cleansing/Transforming

We create the column names, eliminate the first two rows of the table and make some minor fix replacing Australian Open (Jan) and Australian Open (Dec) with Australian Open.

```{r, message = FALSE, warning = FALSE}
# Rename table headings and remove line 1
# Minor fix - naming for the Australian Open tournament (recode - dplyr package)
colnames(tbls_xml) <- c("year", "tournament", "winner","runner-up")
tbls_xml <- tbls_xml %>% slice(2:n())
tbls_xml <- tbls_xml %>% 
  mutate(tournament = str_replace(tournament, "Australian Open (Jan)", "Australian Open")) %>%    mutate(tournament = str_replace(tournament, "Australian Open (Dec)", "Australian Open"))
```

# Plot 

## Total wins

Here we will group by winner and summarise the number of wins. 

```{r, echo = TRUE, message = FALSE, warning = FALSE}
total.wins <- tbls_xml %>% group_by(winner) %>% summarise(wins=n()) %>% arrange(desc(wins))
kable(head(total.wins, 40))
```

We will create a barplot filtered by champions who won at least seven Tennis Grand Slam tournaments. 

* Order the bar plot in descending order -using the function reorder as a solution for this 
* We will be using the ggthemes package (theme and color palettes based on the pander package)

```{r, echo = TRUE, message = FALSE, warning = FALSE}
winners7 <- total.wins %>% filter(wins >= 7)

the_colours = c("#FF4000FF", "#FF8000FF", "#FFFF00FF", "#80FF00FF",
                "#00FF00FF", "#00FF80FF", "#00FFFFFF", "#0080FFFF",
                "#FF00FFFF", "#000000FF", "#0000FFFF")

ggplot(data=winners7, aes(x=reorder(winner, wins), y=wins, fill=wins)) +
  geom_bar(stat='identity') +
  ggtitle("Men's Grand Slam Title Winners") +
  theme_pander() +
  coord_flip() + 
  guides(fill=FALSE) +
  scale_fill_gradientn(colours = the_colours) + 
  theme(legend.title=element_blank()) +
  theme(axis.title.y=element_blank()) +
  theme(axis.title.x=element_blank())
```

## Wins by tournament

Now we will plot total wins by tournament for each winner. 

* Using facet_grid to display each tournament wins
* To compare each champions’ performance for each tournament, we will be grouping by tournament and winner

```{r, echo = TRUE, message = FALSE, warning = FALSE}
total.wins.trn <- tbls_xml %>% filter(winner %in% winners7$winner) %>% group_by(tournament, winner) %>% summarise(wins=n()) %>% arrange(desc(wins)) 

ggplot(data=total.wins.trn, aes(x=reorder(winner, wins), y=wins, fill=wins)) +
  geom_bar(stat='identity') +
  ggtitle("Men's Grand Slam Title Winners") +
  coord_flip() + 
  facet_grid(.~tournament) + 
  guides(fill=FALSE) +
  scale_y_continuous(breaks = seq(0, 10, 2))+
  scale_fill_gradientn(colours = the_colours) + 
  theme(legend.title=element_blank()) +
  theme(axis.title.y=element_blank()) +
  theme(axis.title.x=element_blank()) 
```



[^1]: ESPN Men's Grand Slam Title Winners (http://www.espn.com/tennis/history)