---
title: "Using waffle charts to analyze Spanish general election results, 2016"
author: "Pablo Ugarte"
date: '2017-09-21'
slug: using-waffle-charts-to-analyze-spanish-general-election-results-2016
tags: []
categories: []
---

# Intro

>In this post we wil show how the waffle package can be used plotting 2016 Spanish general elections

If you are interested in learning more about creating charts using the R Waffle package you can read more here:

* Make waffle (square pie) charts in R [^1]
* Infographic-style charts using the R waffle package [^2]

# Libraries

We will first load the packages (you will have to install the packages if you haven't done so before).

```{r, message = FALSE, warning = FALSE}
library(waffle) # Waffle charts
library(stringr) # Manipulating strings
library(tidyverse) # Data wrangling packages gather, spread...
library(readr) # Read docs
library(RColorBrewer) # Colors
library(rvest) # Scraping
library(tidyverse) # Data wrangling
library(mosaic) # Quick exploratory analysis
```

# Scraping

With the rvest package we can scrape the content of the table. 

* We first use the html_nodes() function to select the <table> nodes
* To parse the HTML table data we use html_table(), which would create a list
* We coerce the list to dataframe to be able to work with the data manipulaton

```{r, message = FALSE, warning = FALSE}
webpage <- read_html("https://en.wikipedia.org/wiki/Results_breakdown_of_the_Spanish_general_election,_2016_(Congress)")

# Only parse the table interested
congreso <- webpage %>%
        html_nodes("table") %>%
        .[1] %>%
        html_table(fill = TRUE)

# Coerce list to dataframe
congreso <- as.data.frame(congreso)
```

# Cleansing

If you look at the congreso dataframe variable, the headings are in line 5 and start from column 2. Let's correct this.

```{r, message = FALSE, warning = FALSE}
congreso <- congreso %>% slice(5:15) %>% select(X2,X3,X4,X5)
colnames(congreso) <- c("party", "votes", "percentage", "mp")
congreso <- congreso %>% slice(2:15)
```

Before plotting data, the comma has to be removed so that R will be able to convert it to a numeric variable.  If you don’t remove the comma, you will get an NA for any number value. We will aslo remove the % sign.

```{r, message = FALSE, warning = FALSE}
congreso$votes <- gsub(",","",congreso$votes)
congreso$votes <- as.numeric(congreso$votes)
congreso$percentage <- gsub("%","",congreso$percentage)
congreso$percentage <- as.numeric(congreso$percentage)
```

Next, let's do some quick exploratory analysis on the votes received each political party month using a package called Mosaic.

```{r, message = FALSE, warning = FALSE}
favstats(congreso$votes)
```

The max votes to any one of the political parties is 7,941,236 and the median is 703,048. We are now ready to use the waffle charts to understand more.

# Plot

We will start by giving the colors from the viridis package to the different political parties. Then we can plot.

```{r, message = FALSE, warning = FALSE}
votes <- c(`PP`= 7941236, `PSOE`= 5443846, `UPodemos`= 5087538, `Cs`= 3141570, 
           `ERC–CatSí` = 639652, `CDC` = 483488, `EAJ/PNV` = 287014, `EH Bildu` = 184713, `CCa–PNC` = 81917
, `Others` = 766445) 

waffle(votes/100000, rows=6,
colors=c("#FF8C00", "#00FFFF", "#000000", "#7CFC00", "#006400", 
              "#FFA500", "#FFFF00", "#FFC0CB", "#FF0000", "#0000FF"),
title="Spanish general election results, 2016",
legend_pos="bottom",
xlab="1 seat = 100K people")
```

That's it. The waffle package has many functionalities. If you want to know more click these links or search on the internet for more examples. Nothing more, nothign else!

[^1]: Link1 (https://github.com/hrbrmstr/waffle)
[^2]: Link2 (https://nsaunders.wordpress.com/2017/09/08/infographic-style-charts-using-the-r-waffle-package/)
