---
title: "Electric cars and metal prices"
author: "Pablo Uagarte"
date: '2017-09-14'
slug: electric-cars-and-metal-prices
tags: []
categories: Quandl
---

>In this post we will analyze the evolution of metal prices on the back of the Electric car revolution

# Intro

The electric vehicles industry is creating winners and losers within the world’s biggest metals markets. 

In this regard, Cobalt is essential for lithium-ion batteries powering anything from Tesla Inc.’s cars to Apple Inc.’s iPhones and iPads [^1].

# Libraries

We first install and load the libraries.

```{r, message = FALSE, warning = FALSE}
library(ggplot2) # For amazing charts
library(dplyr) # Data wrangling
library(hrbrthemes) # Aesthetic defaults for ggplot2 charts
library(Quandl) # Get financial data
library(magrittr) # Pipes %>% 
library(tidyverse) # Data wrangling packages
```

# Importing data

We start by getting the prices for Cobalt, Nickel, Aluminium and Copper [^2].

* Normalize = growth from 100
* rdiff = percentage changes,

```{r, message = FALSE, warning = FALSE}
start <- as.Date("2010-01-01")
end <- as.Date("2017-09-14")
metals <- Quandl::Quandl(code = c("LME/PR_CO.1","LME/PR_NI.1","LME/PR_AL.1","LME/PR_CU.1"),
                         start_date=start, end_date=end, transform="normalize")
View(metals)
```

# Cleaning/Transforming

Once we have all the time series prices for these metals:

* Rename specific columns
* Gather each metal within one column variable and also gather the values associated with each metal in a second column variable

```{r, message = FALSE, warning = FALSE}
metals <- metals %>% 
  rename(date = Date, Cobalt = `LME.PR_CO - Cash Buyer`, Nickel = `LME.PR_NI - Cash Buyer`, Aluminum = `LME.PR_AL - Cash Buyer`, Copper = `LME.PR_CU - Cash Buyer`)
metals <- metals %>% gather(metals, price, Cobalt:Copper)
```

# Plot 

```{r, echo = TRUE, message = FALSE, warning = FALSE}
ggplot(metals,aes(x=date,y=price,color=metals)) +
  geom_line(alpha=1) + # Thickness of line
  theme_ipsum(grid_col = "white") + # Gridline white
  labs(
    title='Electric car revolution is shaking up the biggest metals markets',
    subtitle='Prices for a selection of metals from 2010-2017',
    caption='LME prices are reported in US Dollars per metric tonne') +
  scale_y_continuous(labels=scales::comma) + # Commas to make it easier to read large numbers
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") + # Distance between breaks 1 year
  theme(legend.title=element_blank()) + # Remove title for all legends
  theme(plot.background = element_rect(fill = "#cddee7")) 
```

[^1]: Bloomberg (https://www.bloomberg.com/news/articles/2017-08-08/billionaire-who-made-killing-on-cobalt-bets-on-electric-car-fund)
[^2]: Quandle Codes (https://blog.quandl.com/api-for-commodity-data)