---
title: "Joyplots with ggjoy and hrbrthmes"
author: "Pablo Ugarte"
date: '2017-09-12'
slug: joyplots-with-ggjoy-and-hrbrthmes
tags: []
categories: ggjoy
---

>Haven't you run into articles with some data that makes you think, how can visualize the data in a different and possibly better way? 

We will be use Joyplots for this post. Joyplots are partially overlapping line plots that create the impression of a mountain range. They can be quite useful for visualizing changes in distributions over time or space[^1].

The data that we will be looking at is 2016 temperatures in Lincoln, NE[^2][^3].

# Libraries

We first install and load the packages.

```{r, message = FALSE, warning = FALSE}
library(tidyverse) # Data wrangling
library(ggjoy) # calculates density estimates from the provided data and then plots those
library(hrbrthemes) # Aesthetic defaults for ggplot2 charts
library(readr) # Reading data from csv 
library(ggplot2) # Creating charts
```

# Importing data

```{r, message = FALSE, warning = FALSE}
weather.raw <- read_csv("nebraska.csv")
```

# Cleaning/Transforming

Then we cleanse the data that we want to plot, creating the year labels and the factor to help order the years.

```{r, message = FALSE, warning = FALSE}
weather.raw$month <- months(as.Date(weather.raw$CST))
weather.raw$months <- factor(rev(weather.raw$month),levels=rev(unique(weather.raw$month)))

#scales
mins <- min(weather.raw$`Min TemperatureF`)
maxs <- max(weather.raw$`Max TemperatureF`)
View(weather.raw)
```

# Plot

To plot we will be using hrbrthemes::theme_ipsum. With a few lines of code we get a really nice looking joy plot. 

```{r, echo = TRUE, message = FALSE, warning = FALSE}
ggplot(weather.raw,aes(x = `Mean TemperatureF`,y=months,height=..density..)) +
     geom_joy(scale=3) +
      scale_x_continuous(limits = c(mins,maxs)) +
  theme_ipsum(grid=F) +
  theme(axis.title.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.text.y = element_text(angle = 180, hjust = 1)) +
  labs(title='Temperatures in Lincoln NE',
       subtitle='Median temperatures (Fahrenheit) by month for 2016\nData: Original CSV from the Weather Underground')
